---
title: "Assignment 6: Who were the winners?"
author: "Amanda Herbst"
date: 5/13/24
---

At the conclusion of the ASDN project the PIs decided to hand out first, second, and third prizes to the observers who measured the most eggs. Who won? Please use R and dbplyr to answer this question.

Load libraries
```{r, message = FALSE}
library(tidyverse)
library(DBI)
library(dbplyr)
library(duckdb)
```

## Connect to database

```{r}
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir = here::here("week3/database.db"))
```

## Load necessary tables
```{r}
egg_table <- tbl(conn, "Bird_eggs")
nest_table <- tbl(conn, "Bird_nests")
personnel_table <- tbl(conn, "Personnel")
```

## Find the winners
```{r}
winners <- egg_table %>% 
  # join with nest table to get observer ids
  left_join(nest_table, by = c("Nest_ID", "Year", "Site", "Book_page")) %>% 
  # count how many eggs per observer
  group_by(Observer) %>% 
  summarize(total_eggs = n()) %>% 
  # join with personnel table to get full names of observers
  left_join(personnel_table, by = c("Observer"="Abbreviation")) %>% 
  select(Name, total_eggs) %>% 
  # order by numer of eggs
  arrange(-total_eggs) %>% 
  # top 3
  head(3) %>% 
  collect()

winners
```

## `show_query()`

dbplyr was able to complete the above R code with one query using many nested queries. I used `head(3)` to limit to the first 3 rows but the query used `LIMIT 3`.

## Close the connection to the database
```{r}
DBI::dbDisconnect(conn, shutdown=TRUE)
```